Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/Introduction.tex
merging Nirav's changes to intro

Formal verification of cache coherence protocols has traditionally been done
using model checking. That family of analyses significantly restricts the size
as well as complexity of the system that has to be verified; most model-checking
based approaches can not handle more than one level of hierarchy. To the best of
our knowledge, the largest system that has been verified using model checking
has only 12 processors~\cite{Chen:2008:VHC:1559300}, in an era when many
observers are predicting mainstream systems with hundreds of processor cores.

@@

Protocols, post verification, are often described via Term Rewrite Systems
(TRS) \adamc{citation for TRSes?}, which apply equational logic to describe
systems at an interesting point on the spectrum between declarative and
operational methods.  It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a protocol, after its
verification, can be transliterated into Bluespec~\cite{Bluespec:TFRG}, a
hardware description language based on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. Bluespec specifications can be directly implemented into
efficient hardware. In this way one reduces the correctness of the
implementation to the correctness of the protocol, with no need for separate
reasoning about an implementation. While the current Bluespec compiler does not
support some implementation elements that are common in practice (\eg{}
fairness guarantees, atomic actions running over multiple clock cycles), recent
work has provided key insights toward lifting the
restrictions~\cite{Karczmarek}. We believe that Bluespec and its future
evolutions represent a very promising platform for productive implementation of
hardware components, including cache-coherent shared memories, starting from a
description designed to support verification. Our model of the MSI protocol
in Coq is described as an inductively defined set of transitions that is
directly translatable into Guarded Atomic Actions. In fact, our Coq-based model
is a direct transliteration of the MSI protocol specified in Bluespec,
ensuring that the final hardware synthesized from the specification is formally
verified.

>>>>

Formal verification of cache coherence protocols has traditionally been done
using model checking. That family of analyses significantly restricts the size as
well as complexity of the system that has to be verified; most model-checking
based approaches can not handle more than one level of hierarchy. To the best
of our knowledge, the largest system that has been verified using
model checking has only 12 processors, in an era when many observers are
predicting mainstream systems with hundreds of processor cores.
\xxx{citation:
http://www.cs.utah.edu/formal\_verification/pdf/xiaofang\_dissertation.pdf (used
data from their related work must cite)}.

@@

Protocols, post verification, are often described via Term Rewrite Systems
(TRS) \adamc{citation for TRSes?}, which apply equational logic to describe
systems at an interesting point on the spectrum between declarative and
operational methods.  It was shown by Dave
\etal{}~\cite{DNA:CoherenceImplementation} that such a protocol, after its
verification, can be transliterated into Bluespec~\cite{Bluespec:TFRG}, a
hardware description language based on Guarded Atomic Actions, a restriction of
TRS~\cite{Hoe:TCAD}. Bluespec specifications can be directly implemented into
efficient hardware. In this way one reduces the correctness of the
implementation to the correctness of the protocol, with no need for separate
reasoning about an implementation. While the current Bluespec compiler does not
support some implementation elements that are common in practice (\eg{}
fairness guarantees, atomic actions running over multiple clock cycles), recent
work has provided key insights toward lifting the
restrictions~\cite{Karczmarek}. We believe that Bluespec and its future
evolutions represent a very promising platform for productive implementation of
hardware components, including cache-coherent shared memories, starting from a
description designed to support verification.

Our model of the MSI protocol
in Coq is described as an inductively defined set of transitions that is
directly translatable into Guarded Atomic Actions. In fact, our Coq-based model
is a direct transliteration of the MSI protocol specified in Bluespec,
ensuring that the final hardware synthesized from the specification is formally
verified.
